Outline hints

### 2012-07-17 Info/OutlineHints/outline_hints.tex
Added article outline tips.

\usepackage[maxfloats=25]{morefloats}
\newcounter{todoidx}
\usepackage[toc,nonumberlist]{glossaries}
% \setcounter{todoidx}

@@

\title{A GPU Implementation of Capon for an Active Sonar System}

\author{Jo~Inge~Buskenes, %
        Jon~Petter~Aasen, %
        Carl-Inge~Colombo~Nilsen%
\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem All authors are with the Department
of Informatics, University of Oslo, Norway.}% <-this % stops a space

@@

% The paper headers
\markboth{IEEE Journal of Oceanic Engineering}%
{A GPU Implementation of Capon for an Active Sonar System}

@@

% Keywords (normally not used for peer reviews)
\ifPeerReview\else
\begin{IEEEkeywords}
Beamforming, adaptive beamforming, sonar, active, complexity. \todo{not entirely true?}
\end{IEEEkeywords}
\fi}
% \fi

@@

@@

Delays and weights are applied to each of the sensor channels to ensure that signals emanating from a pixel of interest add constructively, while noise and interference from other regions add destructively. This method of adjusting the spatial focus of the antenna is commonly referred to as a beamformer. A

Beamformer algorithms are distinction

The Minimum Variance One of the most well established adaptive

One of the main issues with adaptive beamformers is their computational complexity, and .

Most modern active sonar imaging systems record and process the incoming wavefield through the use of phased arrays. A pixel value in the image is estimated by digitally steering the arrays response to the corresponding region in space. This concept is called beamforming and involves applying a suitable set of delays and weights to each of the sensor channels to cause constrictive interference of the signals coming from the region of interest, and destructive interference of anything else. Ideally, we want a spatial response

If a beamformer only employ a static set of weights and delays, it is referred to as a conventional one. Applying , which  Various weight sets, or windows, may be added, Adaptive beamformers, on the other hand, attempts to find an optimal spatial response that best fits the incoming wavefield.  The most common approach is perform a rough adjustment using a static set of delays, and compute a new set of weights at each time instant.

\todopar{Case:
\begin{itemize}
\item JOE
\item Sonar. Channel count up to 32/64 ish channels?
\item Other articles evaluates Capon image quality, and some have shown that speedups can be achieved by reducing the complexity of the traditional Capon algorithm or by implementing specific versions on a GPU.
\item This article compare a few promising ways to implement Capon in general, in light of general complexity and suitability for implementation on a CPU and GPU.
\end{itemize}
}

\begin{itemize}
\item As briefly as possible, introduce beamforming.
\item Adaptive beamformer's potential lies in its ability to suppress interference power
\item Why adaptive beamformers struggle in active sonar systems. Correlated noise, robustification kills the adaptive potential. Quite computationally intensive. Constraints must be applied in one way or another - parameters must be tuned.
\item The three main ways to implement Capon \cite{Capon1969} is \todo{not entirely true?}
\begin{itemize}
\item Traditional way: Build $\eR$ and solve $\w = \frac{\Ri\a}{\a\H\Ri\a}$. 
\item Iterative methods: Woodbury, Conjugated gradients
\item Beamspace
\item LCA (or other methods, as reference)
\end{itemize}
\item Choice of robustification parameters $L,M,\Delta$ largely impacts what the ``optimal'' way of implementing Capon 
\item In most imaging applications 
Outline:
\begin{itemize}
\item Speeding up: Build $\eR$ and solve $\w = \frac{\Ri\a}{\a\H\Ri\a}$. 
\item Iterative methods: Woodbury (Sherman Morrison), Conjugated gradients - show performance of GPU performance.
\item Woodbury: Must take the average over many samples in space and time - lots of (small) outerproducts. Can be used if not a lot of averaging is required.
\item Beamspace
\item LCA (or other methods, as reference)
\end{itemize}
\end{itemize}

@@

\begin{align}
z[n] = \sumb{m=0}{M-1} w_m[n]^*x_m[n-\Delta_m] = \w\H[n]\Xd[n]
\end{align}
where
\begin{align}
\w[n] = \bmat{w_0[n]\\w_1[n]\\\vdots\\w_{M-1}[n]} \quad \text{and} \quad\Xd[n] = \bmat{x_0[n]\\x_1[n]\\\vdots\\x_{M-1}[n]}.
\end{align}

Minimum variance distortionless resonse \cite{Capon1969}
\begin{align}
\underset{\w[n]}{\argmin}\, E\{ |z[n]|^2 \} = \underset{\w[n]}{\argmin}\, \w\H[n]\R[n]\w[n], 
\end{align}
subject to
\begin{align}
\w\H[n]\a = 1,
\end{align}
where
\begin{align}
\R[n] = E\{ \x\x\H \}.
\end{align}
If $\eR$ is the estimate of $\R$, the solution to the MVDR beamformer is
\begin{gather*}
\vec w[n] = \frac{\hat{\mat R}\,\!^{-1}[n]\vec a}{\vec a\H\hat{\mat R}\,\!^{-1}[n]\vec a} = \frac{\text{\raisebox{1.9pt}{$\vec\chi$}}[n]}{\vec a\H\text{\raisebox{1.9pt}{$\vec\chi$}}[n]}
\end{gather*}
A robust estimate of $\eR$ is found by
\begin{itemize}
\item Spatial averaging to decorrelate noise from signal. \emph{Subarray averaging} used here.
\item Temporal averaging to ensure valid speckle statistics.
\item \emph{Diagonal loading} to ensure  numerical stability prior to inversion.
\item Choice of robustification parameters \cite{Synnevag2007}
\end{itemize}

Here $\vec a$ is a \emph{steering vector} that if set to $\vec 1$ steers towards broadside, and $\hat{\mat R}$ is an estimate of the spatial covariance matrix, often computed as
\begin{gather*}
r_{ij}[n] = \frac{1}{K(2N_{\text{avg}}+1)}\sumb{k=0}{K-1} \sumb{n'=n-N_{\text{avg}}}{n+N_{\text{avg}}} x_{i+k}[n']\,x_{j+k}[n'],
\end{gather*}
where $r_{ij}$ is the $(i,j)$'th element of the covariance matrix, $K$ is the number of subarrays, $2N_{\text{avg}}+1$ is the number of samples to perform time averaging over, and $\vec x_i[n']$ is the data recorded by the $i$'th sensor at sample $n'$.

\emph{Diagonal loading} chosen as \cite{Synnevag2007}
\begin{align}
\varepsilon = \Delta \cdot \tr\{\eR[n]\}
\end{align}

\subsection{Traditional approach}

Let us first consider the \emph{computation of $\hat{\mat R}$}. Notice that when implementing the formula directly a lot of summations and multiplications will be repeated unnecessary. To illustrate this, consider for instance a case with $M=6$ sensors that are grouped into subarrays of length $L=3$ where all but one element overlap. Let us further assume that we need to average over 7 time samples, i.e. that $N_{\text{avg}} = 3$.

These procedures are commutative, so the order in which we perform them does not matter. But for the sake of this example let us start by performing the time averaging. Assuming that $\vec x_i$ and $\vec x_j$ have a zero mean, a time averaged covariance coefficient $r_{ij}$ in $\hat{\mat R}$ may be obtained by (omitting the normalisation):
\begin{align*}
r_{ij}[n] &= \sumb{n'=n-N_{\text{avg}}}{n+N_{\text{avg}}} x_i[n']\,x_j[n'],
\end{align*}
which can be easily rewritten in a sliding window fashion
\begin{align*}
r_{ij}[n] &= r_{ij}[n-1] - x_i[n\text{-}N_\text{avg}\text{-}1]\,x_j[n\text{-}N_\text{avg}\text{-}1]
                         + x_i[n\text{+}N_\text{avg}]\,x_j[n\text{+}N_\text{avg}].
\end{align*}
Hence, whenever $N_\text{avg}>1$ the sliding window approach will demand less instructions at the expense of a slightly higher memory consumption. The next thing to realise is that not all $r_{ij}[n]$ need to be computed. Since $r_{ji} = r^*_{ij}$, and the subarrays only span $L$ elements, we only compute an upper triangular band of R (marked \Green{green})
\begin{figure}[H]
% \begin{narrow}{-3.5cm}{-3.5cm}
\includegraphics[width=\linewidth]{gfx/capon_build_R_full.eps}
\caption{Building $\hat{\mat R}$: Symmetry and element coverage}\label{buildingRfull}
% \end{narrow}
\end{figure}
The final step is to average all the subarray matrices to form $\hat{\mat R}$ by 
{\renewcommand*{\arraystretch}{1.8}
\begin{align*}
\hat{\mat R}_\text{sub} &= {\color{Blue}\mat R_{\text{sub},1} + \mat R_{\text{sub},2} + \mat R_{\text{sub},3} + \mat R_{\text{sub},4}} \\
&= \left[\;
\begin{matrix}
\color{Green}r_{00}+r_{11}+r_{22} & \color{Red}r_{01}+r_{12}+r_{23} & r_{02}+r_{13}+r_{24} \\
                                       & \color{Green}r_{11}+r_{22}+r_{33} & \color{Red}r_{12}+r_{23}+r_{34} \\
                                       &                                        & \color{Green}r_{22}+r_{33}+r_{44}
\end{matrix}
\;\right]
\end{align*}}%
Note that consecutive elements along the diagonals overlap on all but one covariance coefficient, hence a sliding window may also be used here. The subarray size $L$ is usually quite a bit larger than the time averaging window, so there is more to gain from implementing the sliding window here.\\\\

{\bf Method complexity:}\\
\begin{tabular}{l l}
No optimizations            & $2M^2\,(2\Navg+1)$ \\
Banded symmetric            & $L(2M-L+1)\,(2\Navg+1)$ \\
Sliding, banded, symmetric  & $1.5L(2M-L+1)$ \\
\end{tabular}\\\\

\subsubsection{Computing $\text{\raisebox{1.9pt}{$\vec\chi$}} = \hat{\mat R}\,\!^{-1}\vec a$}

There are two ways of finding $\text{\raisebox{1.9pt}{$\vec\chi$}} = \hat{\mat R}\,\!^{-1}\vec a$, either by solving the equation directly or by inverting $\hat{\mat R}$ first. Solving the equation is less computationally intensive than inverting $\hat{\mat R}$, and seems like the obvious choice. However, with iterative methods for updating $\eRi$ for each range pixel the situation is changed...

\subsection{Iterative methods}

\begin{itemize}
\item Woodbury
\item Conjugated gradients
\end{itemize}

\subsection{Beamspace}

\begin{itemize}
\item Some minimum of background, and good references.
\end{itemize}

\section{Results/Discussion}

\begin{itemize}
\item Compare computational complexity vs. memory constraints for the various methods. Supporting plots.
\item Relate this to whether the favourable architecture is a CPU or a GPU \todo{check ``when to gpu article''}
\item Run benchmarks on some sonar dataset(s)
\end{itemize}

@@

\begin{itemize}
\item When to use what.
\end{itemize}

% \ \\
% \IEEEPARstart{T}{o} form images from a modern phased array sonar system the received wavefield is usually recorded, and then postprocessed by a digital beamformer. The beamformer applies delays and weights to the sensor channels, the beamformer adjusts the arrays spatial response to focus at one pixel at a time.  such that signals emanating from regions of interest add constructively, while ensuring that noise and interference from other angles do not. 
% 
% The imaging capabilities of a modern phased array sonar system depend on physical attributes such element response and array geometry, the transmitted signal, as well as the beamforming method being used on transmission and reception. Beamforming is the concept of applying delays and weights to the sensors channels to steer the arrays response to points of interest.

% 
% 
% Outline:
% \begin{itemize}
% \item Capon's resonse when applying robustification
% \item Choice of window functions makes little difference.
% \item Steering and mainlobewidths have outer bounds.
% \item Beamspace?
% \item Chosen window plots - what may they tell us? Variance intensity values when using various windows.
% \item Assymmetric windows?
% \end{itemize}
% 
% 
% \begin{align}
% z[n] = \sumb{m=0}{M-1} w_m[n]^*x_m[n-\Delta_m] = \w\H[n]\x[n-\Delta_m]
% \end{align}
% 
% 
% \section{Methods}
% 
% Basically, we are working with a practical implementation of the Capon beamformer that computes a set of weights $\vec w$ for every single sample $n$ by solving:
% \begin{gather*}
% \vec w[n] = \frac{\hat{\mat R}\,\!^{-1}[n]\vec a}{\vec a\H\hat{\mat R}\,\!^{-1}[n]\vec a} = \frac{\text{\raisebox{1.9pt}{$\vec\chi$}}[n]}{\vec a\H\text{\raisebox{1.9pt}{$\vec\chi$}}[n]}
% \end{gather*}%
% where
% % \newcommand\X{\text{\raisebox{2pt}{$\vec\chi$}}}
% \begin{gather*}
% \text{\raisebox{1.9pt}{$\vec\chi$}} = \hat{\mat R}\,\!^{-1}\vec a \qquad\qquad\text{is the solution to}\qquad\qquad \hat{\mat R}\text{\raisebox{1.9pt}{$\vec\chi$}} = \vec a.
% \end{gather*}

% 
% , and maximum suppression of while ensuring that the beamformer digitally  before each of the pixels are estimated one at a time. The resolution and contrast of such a system will depend on the systems spatial response, which ideally should be narrow  be very sharp in the desired direction its ability to achieve  fundamental principle of forming a sonar image is to record the received wavefield, 
% 
% image quality of a phased array sonar imaging systems depend on  the choice of weights to apply to each of the sensors are crucial. 
% 
% A modern phased array imaging system may be thought of as a spatial filter. To achieve the best possible performance, the 
% 
% resolution and contrast 
% 
% Adaptive beamformers have only recently been introduced in active sonar imaging. For a while they were considered unsuited for this purpose because the backscattered signal is largely correlated with the 
% 
% 
%

%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation Results}
%\label{fig_sim}
%\end{figure}

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command, the
% \label for the overall figure must come after \caption.
% \hfil must be used as a separator to get equal spacing.
% The subfigure.sty package works much the same way, except \subfigure is
% used instead of \subfloat.
%
% \begin{figure*}[!t]
% \centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
% \label{fig_first_case}}
% \hfil
% \subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
% \label{fig_second_case}}}
% \caption{Simulation results}
% \label{fig_sim}
% \end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.

% An example of a floating table. Note that, for IEEE style tables, the 
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that IEEE does not put floats in the very first column - or typically
% anywhere on the first page for that matter. Also, in-text middle ("here")
% positioning is not used. Most IEEE journals use top floats exclusively.
% However, Computer Society journals sometimes do use bottom floats - bear
% this in mind when choosing appropriate optional arguments for the
% figure/table environments.
% Note that, LaTeX2e, unlike IEEE journals, places footnotes above bottom
% floats. This can be corrected via the \fnbelowfloat command of the
% stfloats package.

\section{Conclusion}
The conclusion goes here.

% if have a single appendix:
%\appendix[Proof of the Zonklar Equations]
% or
%\appendix  % for no appendix heading
% do not use \section anymore after \appendix, only \section*
% is possibly needed

% use appendices with more than one appendix
% then use \section to start each appendix
% you must declare a \section before using any
% \subsection or using \label (\appendices by itself
% starts a section numbered zero.)
%

@@

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

\bibliographystyle{IEEEtran}
\bibliography{../../Library/library}

% End up doing this:
% \begin{thebibliography}{1}
% 
% \bibitem{IEEEhowto:kopka}
% H.~Kopka and P.~W. Daly, \emph{A Guide to {\LaTeX}}, 3rd~ed.\hskip 1em plus
%   0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
% 
% \end{thebibliography}

% biography section
% 
% If you have an EPS/PDF photo (graphicx package needed) extra braces are
% needed around the contents of the optional argument to biography to prevent
% the LaTeX parser from getting confused when it sees the complicated
% \includegraphics command within an optional argument. (You could create
% your own custom macro containing the \includegraphics command to make things
% simpler here.)
%\begin{biography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
% or if you just want to reserve a space for a photo:

\vfill

\begin{IEEEbiography}{Noname dude}
He doesn't exist.
\end{IEEEbiography}

% if you will not have a photo at all:
\begin{IEEEbiographynophoto}{Jo Inge Buskenes}
Not in at the moment.
\end{IEEEbiographynophoto}

%\newpage

% \begin{IEEEbiographynophoto}{Jo Inge Buskenes}
% Not in at the moment.
% \end{IEEEbiographynophoto}

%\vfill

% Can be used to pull up biographies so that the bottom of the last one
% is flush with the other column.
%\enlargethispage{-5in}

\markboth{}%
{}
% \input{gfx/gfx_input}

>>>>

\usepackage[maxfloats=25]{morefloats}
\newcounter{todoidx}
% \setcounter{todoidx}

@@

\title{Outline hints}

\author{Authors%
\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem All authors are with the Department
of Informatics, University of Oslo, Norway.}% <-this % stops a space

@@

% The paper headers
\markboth{Some journal}%
{Outline hints}

@@

% Keywords (normally not used for peer reviews)
\ifPeerReview\else
\begin{IEEEkeywords}
Outline hints
\end{IEEEkeywords}
\fi}
% \fi

@@

\section{Prerequisites}

\begin{itemize}
\item Research articles. Take note of journals aim for these. Perhaps take impact factor into account.
\item For a given journal, make note of typical article and section lengths.
\item Notice language and outline of such articles as well.
\end{itemize}

IMRAD style:

\section{Abstract}

\begin{itemize}
\item Past tense.
\end{itemize}

@@

\begin{itemize}
\item Establish audience, and outline what they need to know.
\item Purpose: What the audience need to know in order to understand the findings of the paper.
\item Reason: Why is the research in this field necessary, and where does this individual contribution fit in?
\item Active voice.
\item Present tense, as focus will be put on the problem at hand, and respect should be paid to established knowledge.
\end{itemize}

\begin{enumerate}
\item Nature and scope of the problem, with all possible clarity. Motivate!!! Why the subject was chosen, why it is important.
\item Brief review of present litterature for orientation purposes. 
\item Method of investigation? Optionally, why was it chosen? Mediate our way of thinking!
\item Principal results. Some overlap with abstract ok.
\item Principal conclustions suggested by the results (don't keep the reader in suspense). Some overlap with abstract ok.
\end{enumerate}
\begin{itemize}
\item Funnel effect desirable: Importance of overall topic, summarize knowledge about an aspect of it, identify an uresolved question of that aspect, say how paper answers that question.
\end{itemize}

@@

\begin{itemize}
\item Describe and defend method sufficently to allow experiment to be repeated.
\item Even though most readers will skip it since method is stated in introduction, results will not be taken seriously if method is insufficiently documented.
\item Use subheadings to group methods, try to make them match the ones used in Results.
\item Like a cookbook: ``How'', and ``how much'' should be precisely answered.
\item Statistical knowledge - of established knowledge - should not be included. If deemed important, cite relevant litterature.
\item When referencing, state the findings, but omit the reasons behind the findings.
\item Passive voice appropriate, of use of ``we''.
\item Past tense.
\end{itemize}

\section{Results}

\begin{itemize}
\item Digest results until only the vital pieces of information remains. Make what is important stand out.
\item Drop variables that are neglible, and be sure to mention it.
\item If negative aspects are meaningful, include them.
\item Strive for clarity! This is the new piece of information offered to the world. Make sure it sees it.
\item Results define everything. The introduction and methods explain why and how they come into existence, discusstion explains what they mean.
\item Avoid redundancy. State data in figures/tables, not in text.
\item Although dominated by passive voice, add some active voice to the mix too.
\item Past tense on own results.
\end{itemize}

\begin{enumerate}
\item Overall description of the experiment, providing ``big picture'' without repeating technicalities of methods section.
\item Results, with crystal clarity.
\end{enumerate}

\section{Discussion}

\begin{enumerate}
\item Present principles, relationships, and generalizations shown by the Results. Discuss, not recapitulate.
\item Point out exceptions, lack of correlation, or unsettled points.
\item Show how results and interpretations agree/contrast with previously published work, and with own expectations.
\item Discuss theoretical implications, as well as possible practical applications. This will largely affect how significant the article will be percieved.
\item State conclusions are clearly as possible.
\item Summarize evidence for each conclusion. Beware of causality leaps!
\item Discussion pairs Introduction. Be sure to answer any question asked there. Don't recapitulate, but extend from where introduction left off.
\item Be sure to emphasize that this is a ``small truth'', not the ``whole truth''.
\item Active voice.
\end{enumerate}

@@

\begin{enumerate}
\item Take a step back. Summarize experiment, relating it to the big picture.
\item Discuss further work
\item Did the results support / reject the hypothesis stated in the introduction?
\item If the used should remember just one sentence from your article, state it here!
\end{enumerate}

\section{Revision}

\begin{itemize}
\item Information selection wise?
\item Information accurate?
\item Consistency?
\item Organization logical?
\item Clear wording?
\item Points stated brielfy, simply, directly?
\item Grammar, spelling, word use, correct?
\item Check figures and tables for errors/consistency.
\item IEEE compliant?
\end{itemize}

@@

% \bibliographystyle{IEEEtran}
% \bibliography{../../Library/library}
